Bioconductor workflow for single-cell RNA sequencing: Normalization, dimensionality reduction, clustering, and lineage inference

### 2017-12-05 workflow/workflow.Rmd
comments

```{r options, echo=FALSE, results="hide",message=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
if(packageVersion("zinbwave")<"0.99.6") stop("must have current devel version to avoid bugs")
if(packageVersion("clusterExperiment")<"1.5.0-9007") stop("must have current devel version to avoid bugs")
if(packageVersion("scone")<"1.1.2") stop("must have current devel version to avoid bugs")

@@

EP: Why do we create 'core' rather than just replace 'dataObj'. I agree, I think that it is unnecessary. I changed 'core' to dataObj.

@@

```{r zinb}
print(system.time(dataObj <- zinbwave(dataObj, K = 50, X = "~ Batch", residuals = TRUE, normalizedValues = TRUE)))
```

@@

```{r addW}
reducedDim(dataObj,"zinbwave50")<-W
```

@@

```{r rsec_50}
print(system.time(dataObj <- RSEC(dataObj, k0s = 4:15, alphas = c(0.1),
                                betas = 0.8, reduceMethod="zinbwave50",
                                clusterFunction = "hierarchical01", minSizes=1,
                                ncores = NCORES, isCount=FALSE,
                                dendroReduce="zinbwave50",
                                subsampleArgs = list(resamp.num=100,
                                                     clusterFunction="kmeans",
                                                     clusterArgs=list(nstart=10)),
                                verbose=TRUE,
                                combineProportion = 0.7,
                                mergeMethod = "none", random.seed=424242,
                                combineMinSize = 10)))
```

@@

@@

@@

@@

EP: publishedClusters is saved as a cluster, so don't need to set it up. Now normalized should be default assay, so not need 'visualizeData'

@@

```{r mdsWce, fig.cap="RSEC: MDS of the low-dimensional matrix W, where each point represents a cell and cells are color-coded by RSEC clustering."}
plotDimReduce(dataObj,whichCluster="primary",reduceMethod="zinbwave50",pch=20,
	xlab = "Component1", ylab = "Component2",legendTitle="Sample",main=""
)
```

@@

@@

```{r}
pseudoCe<-dataObj[,!primaryClusterNamed(dataObj) %in% c("-1", "c4")]
X <- reducedDim(pseudoCe,type="zinbwave50")
mds <- cmdscale(dist(X), eig = TRUE, k = 4)
X <- mds$points
lineages <- getLineages(X, clusterLabels = primaryClusterNamed(pseudoCe), start.clus = "c1")
```

EP: created a pseudoCe to subset to the samples for the lineage. Simplifies code enormously and as long as we don't change it, doesn't use additional memory, I think.

>>>>

```{r options, echo=FALSE, results="hide",message=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
if(packageVersion("zinbwave")<"1.1.0") stop("must have current master version to avoid bugs")
if(packageVersion("clusterExperiment")<"1.5.0-9007") stop("must have current devel version to avoid bugs")
if(packageVersion("scone")<"1.2.0") stop("must have current master version to avoid bugs")

@@

EP: Why do we create 'core' rather than just replace 'dataObj'.   
FP: I agree, I think that it is unnecessary. I changed 'core' to dataObj.

@@

```{r zinb,eval=FALSE}
print(system.time(dataObj <- zinbwave(dataObj, K = 50, X = "~ Batch", residuals = TRUE, normalizedValues = TRUE)))
```

```{r}
#save(dataObj, file= 'dataObj.rda')
load('dataObj.rda')
```

@@

```{r addW}
SingleCellExperiment::reducedDim(dataObj,"zinbwave50")<-W
```

@@

```{r rsec_50,eval=FALSE}
print(system.time(dataObj <- RSEC(dataObj, k0s = 4:15, alphas = c(0.1),
                                betas = 0.8, reduceMethod="zinbwave50",
                                clusterFunction = "hierarchical01", minSizes=1,
                                ncores = NCORES, isCount=FALSE,
                                dendroReduce="zinbwave50",
                                subsampleArgs = list(resamp.num=100,
                                                     clusterFunction="kmeans",
                                                     clusterArgs=list(nstart=10)),
                                verbose=TRUE,
                                combineProportion = 0.7,
                                mergeMethod = "none", random.seed=424242,
                                combineMinSize = 10)))
```

```{r}
#save(dataObj, file = 'dataObj_afterRSEC.rda')
load('dataObj_afterRSEC.rda')
```

@@

FP: Ordering does not seem right.

@@

FP: we do not get exactly the same number of cells in each cluster. It looks similar though.

@@

FP: compared to the previous version, it seems that there is a lot of black. Would the black coming from the contour of the rectangles in the bars?

@@

EP: publishedClusters is saved as a cluster, so don't need to set it up. Now normalized should be default assay, so not need 'visualizeData'.

FP: publishedClusters does not appear in the bars.

@@

```{r mdsWce, fig.cap="RSEC: MDS of the low-dimensional matrix W, where each point represents a cell and cells are color-coded by RSEC clustering."}
plotDimReduce(dataObj,whichCluster="primary",reduceMethod="zinbwave50",pch=20,
	xlab = "Component1", ylab = "Component2",legendTitle="Sample",main=""
)
```
FP: I get `"reduceMethod" is not a graphical parameter`. I think this plot is different from the previous version.

@@

FP: this plot is fairly different from the previous version. We should see for one cluster that the gene expression for Reg3g is much high that for the other clusters.

@@

```{r}
pseudoCe <- dataObj[,!primaryClusterNamed(dataObj) %in% c("-1", "c4")]
X <- reducedDim(pseudoCe,type="zinbwave50")
mds <- cmdscale(dist(X), eig = TRUE, k = 4)
X <- mds$points
lineages <- getLineages(X, clusterLabels = primaryClusterNamed(pseudoCe), start.clus = "c1")
```

EP: created a pseudoCe to subset to the samples for the lineage. Simplifies code enormously and as long as we don't change it, doesn't use additional memory, I think.  
FP: I agree. I get an error though when I run line 501:
`Error in .BasicFunsList[[f]] : no such index at level 1`.

FP: I stooped here.
